---
title: "10_import"
output: html_notebook
---

# Load the libraries

The libraries I need 

```{r import_libraries, warning=FALSE}
# Install the package "wru" if you need
# install.packages("wru")
library(readr)
library(janitor)
library(assertr)
library(tidyverse)
library(glue)
library(scales)
library(wru)
```

# Load the data set

Load the data and clean the feature names

```{r import_data}
train <- read_csv("data/train.csv") %>% 
  clean_names()
```

# Initially check the data

```{r init_check}
dim(train)
str(train)
```

From the message above, we can see that there are 891 observations and we have 12 different features.

Here is the data dictionary

Variables | description
---|---
**passenger_id** | An arbitrary ID assigned to each passenger
**survived** | The suvival status of the passenfer (1 = Survived, 0 = Did not survive)
**pclass** | The passenger’s class number (1 = 1st; 2 = 2nd; 3 = 3rd)
**name** | The passenger’s first name, last name, and title
**sex** | The passenger’s gender
**age** | The passenger’s age (in years)
**sib_sp** | The number of siblings/spouses aboard the ship with the passenger
**parch** | The number of parents/children aboard the ship with the passenger
**ticket** | The passenger’s ticket number
**fare** | How much the passenger paid to board the ship
**cabin** | The passenger’s cabin number
**embarked** | Where the passenger embarked from (C = Cherbourg; Q = Queenstown; S = Southampton)

# Test the factor features

From the overview of our dataset, we should test all the string feature which should be factors in this part.

## Test the "survived" variable

Check if all of the values are 0 or 1.

```{r test_survived}
train %>% 
  assert(in_set(0, 1), survived)
```

Great!  We do not have outliers for the "survived" feature, all of them are 0 or 1.

## Test the "pclass" variable

Check if all of the values are 1 or 2 or 3.

```{r test_pclass}
train %>% 
  assert(in_set(1, 2, 3), pclass)
```

Great!  We do not have outliers, all of them are 1 or 2 or 3.

## Test the "sex" variable

Check if all of the values are "male" or "female".

```{r test_sex}
train %>% 
  assert(in_set("male","female"), sex)
```

Great!  We do not have outliers for the "sex" feature, all of the values are "male" or "female".

## Test the "embarked" variable

Check if all of the values are "C" or "Q" or "S".

```{r test_embarked}
train %>% 
  assert(in_set("C","Q", "S"), embarked)
```

Great!  We do not have outliers for the "embarked" feature, all of the values are "C" or "Q" or "S".

# Test the numeric variables

For this part, we shoud test every numeric variables.

## Test the "age" variable

Check if all the values are between 0 and 200.

```{r test_age}
train %>% 
  assert(within_bounds(0,200), age)
```

Great!  We do not have outliers for the "age" feature, all the values are between 0 and 200.

## Test the "sib_sp" variable

Check if all the values are bigger than 0.

```{r test_sibsp}
train %>% 
  assert(within_bounds(0,Inf), sib_sp)
```

Great!  We do not have outliers for the "sib_sp" feature, all the values are bigger than 0.

## Test the "parch" variable

Check if all the values are bigger than 0.

```{r test_parch}
train %>% 
  assert(within_bounds(0,Inf), parch)
```

Great!  We do not have outliers for the "parch" feature, all the values are bigger than 0.

## Test the "fare" variable

Check if all the values are bigger than 0.

```{r test_fare}
train %>% 
  assert(within_bounds(0,Inf), fare)
```

Great!  We do not have outliers for the "fare" feature, all the values are bigger than 0.

# Data cleaning

From the previous section we know that we don't have outliers for our features, for this part, we are going to remove the NA values and convert the data columns.

## Check the NA values

Check all the NA values

```{r check_all_NA}
colSums(is.na(train))
```

We can see that we have 177 NA values in "age", 687 in "cabin", 2 in "embarked".

## Remove the NA values.

For the "cabin" feature, the count of NA values are more than half, so we decide to remove this feature, for "age" and "embarked", the missing values do not take a really huge proportion so we just remove those NA values.

```{r remove_NA}
data_complete <- train %>% 
  select(everything(),-cabin) %>% 
  na.omit()
data_complete
```

## Convert data columns.

For our data, we should change the "survived", "pclass", "sex", "embarked" into factor variables.

```{r convert_columns}
clean_data <- data_complete %>% 
  mutate(survived = as.factor(survived),
         pclass = as.factor(pclass),
         sex = as.factor(sex),
         embarked = as.factor(embarked))
```

## Check the clean data

```{r check_clean_data}
str(clean_data)
```

The data is clean now.